---
title: "Food loss data and figures"
author: "Johannes Piipponen"
date: "Spring 2023"
output: html_document
---

Code written by Johannes Piipponen



Feel free to edit -- The code now includes multiple sections that involve repetitive copy-pasting and modification of similar elements. Instead of this approach, a more efficient solution could have been achieved by utilizing (stronger) functions, among other techniques.

However, given the circumstances, this was the fastest way for me to complete the task and the easiest to debug.

So, please be patient and allow the code to run. It may take some time, but it should work!
- BR Johannes Piipponen

# Load relevant packages

```{r}
# Define the required packages
packages <- c("tidyverse", "scico", "tmap", "sf", "terra", "here")

# Check if the packages are already installed
not_installed <- packages[!(packages %in% installed.packages()[,"Package"])]

# Install the packages if they are not already installed
if (length(not_installed)) {
  install.packages(not_installed)
}

# Load the required packages
lapply(packages, require, character.only = TRUE)  # You can also use "library" instead of "require"

```


# Load some other functions, data etc

```{r}
# Create a template raster from scratch (needed later)
template_rast_5arcmin <- 
  rast(nrows = 2160, ncols = 4320, crs = "EPSG:4326")

# ---------------- Load polygons that contain geometries for different countries
# Originally loaded from https://www.naturalearthdata.com/downloads/
# Simplified using ms_simplify and modified based on the code below

# adm_10m <- here("Data", "Input", "ne_10m_admin_0_countries.shp") |>
#   read_sf()
# adm10_simple <- ms_simplify(adm_10m)  # 203 rows -- was 258

# Change adm10 iso codes to fao codes
# adm10_simple_faoadded <- adm10_simple %>%
#   dplyr::select(ADMIN, NAME, SOVEREIGNT, ISO_A3_EH, REGION_UN) %>%
#   as.data.frame() %>%
#   st_drop_geometry()
# 
# adm10_simple_faoadded <- adm10_simple_faoadded %>%
#   mutate(
#     fao_from_iso3eh = countrycode(.$ISO_A3_EH, origin = "iso3c", destination = "fao"),
#     fao_from_SOVEREIGNT = countrycode(.$SOVEREIGNT, origin = "country.name",  destination = "fao")
#   )
# ## Warns that some are missing. However, either fao_from_iso3eh or fao_from_SOVEREIGNT includes most
# 
# ## Give Somaliland FAO code of Somalia
# adm10_simple_faoadded <- adm10_simple_faoadded %>%
#   rows_update(
#     .,
#     tibble(
#       SOVEREIGNT = "Somaliland",
#       fao_from_SOVEREIGNT = (
#         filter(adm10_simple_faoadded, SOVEREIGNT == "Somalia") %>%
#         pull(fao_from_SOVEREIGNT)
#       )
#     )
#   )
# 
# ## Combine columns fao_from_iso3eh and fao_from_SOVEREIGNT.
# ## If any of the columns has a value, this value will be the FAO_ID
# adm10_simple_faoadded <- adm10_simple_faoadded %>%
#   mutate(FAO_ID = coalesce(fao_from_iso3eh, fao_from_SOVEREIGNT)) %>%
#   filter(SOVEREIGNT != "Antarctica")  ## Drop Antarctica
# 
# adm10_simple_faoadded <- st_as_sf(adm10_simple_faoadded)

# Load preprocessed country polygons with FAO ID
adm_10m_fao_id_simple <- st_read("adm_10m_fao_id_simple.gpkg")

# Project adm_10m_fao_id_simple to Robinson projection
adm_10m_fao_id_simple_rob <- adm_10m_fao_id_simple %>% 
  st_transform(., crs = "ESRI:54030")

# Convert the country polygon to raster format
cntry_raster <- 
  rasterize(vect(adm_10m_fao_id_simple), template_rast_5arcmin, 
            field = "FAO_ID")

# Function that makes plotting a bit easier
# This function takes arguments:
# r_index: the raster to be plotted
# index_label: represents the title of the plot
# index_main_title: represents the main title of the plot
# colorpal: the color palette we select
# breakvals: the break values for the color palette
# breaknames: the names for the breaks
# color_midpoint: the midpoint value for the color palette
# tocrs: we often want to use the Robinson projection when plotting something
create_index_map <- function(r_index, index_label, index_main_title, colorpal, breakvals, breaknames = NULL, color_midpoint = NULL, tocrs = NA) {
  if (!is.na(tocrs)) {
    r_index <- project(r_index, tocrs, mask = TRUE)
  }
  index_map <- tm_shape(r_index) +
    tm_raster(
      palette = colorpal,
      breaks = breakvals,
      labels = breaknames,
      title = index_label,
      midpoint = color_midpoint,
      legend.is.portrait = FALSE
    ) +
    tm_layout(
      main.title = index_main_title,
      main.title.position = "center",
      main.title.size = 0.8,
      legend.bg.color = TRUE,
      legend.outside = TRUE,
      legend.title.size = 1,
      legend.text.size = 1,
      legend.outside.size = 0.2,
      legend.outside.position = "bottom",
      frame = FALSE
    ) +
    tm_shape(adm_10m_fao_id_simple_rob) +
    tm_borders(col = NA, lwd = 0.5)
  
  return (index_map)
}

# ----------------------------------------------------------- Get Daniel's data
daniel_original <- read.csv("data_johannes.csv")


# Combine geometries and food loss data
country_data <- adm_10m_fao_id_simple %>% 
  left_join(daniel_original, by = c("ISO_3166_1" = "cntry_code"))

pal <- scico(n = 10, palette = "nuuk", direction = -1)

```



# Figure 1

Average food loss data distribution based on different commodities across food supply stages: wheat (a), maize (b), rice (c), other cereals & pulses (d), fruits & vegetables (e), and roots, tubers, & oil-bearing crops (f).



```{r}
(plt_wheat <- 
    rasterize(country_data %>% filter(basketItem2 == "wheat") %>% 
                group_by(FAO_ID, country) %>% 
                summarise(foodloss_mean = mean(loss_percentage1)) ,
              template_rast_5arcmin, "foodloss_mean")  %>% 
    create_index_map(., tocrs = "ESRI:54030",
                     index_main_title = "(a) Wheat",
                     index_label = "Average food loss (%)",
                     breakvals = c(0,2,4,6,8,10,15,20,Inf),
                     breaknames = c("<2", "2-4", "4-6","6-8", "8-10", "10-15", "15-20", ">20"),
                     colorpal = pal))





(plt_maize <- 
    rasterize(country_data %>% filter(basketItem2 == "maize") %>% 
                group_by(FAO_ID, country) %>% 
                summarise(foodloss_mean = mean(loss_percentage1)) ,
              template_rast_5arcmin, "foodloss_mean")  %>% 
    create_index_map(., tocrs = "ESRI:54030",
                     index_main_title = "(b) Maize",
                     index_label = "Average food loss (%)",
                     breakvals = c(0,2,4,6,8,10,15,20,Inf),
                     breaknames = c("<2", "2-4", "4-6","6-8", "8-10", "10-15", "15-20", ">20"),
                     colorpal = pal))



(plt_rice <- 
    rasterize(country_data %>% filter(basketItem2 == "rice") %>% 
                group_by(FAO_ID, country) %>% 
                summarise(foodloss_mean = mean(loss_percentage1)) ,
              template_rast_5arcmin, "foodloss_mean")  %>% 
    create_index_map(., tocrs = "ESRI:54030",
                     index_main_title = "(c) Rice",
                     index_label = "Average food loss (%)",
                     breakvals = c(0,2,4,6,8,10,15,20,Inf),
                     breaknames = c("<2", "2-4", "4-6","6-8", "8-10", "10-15", "15-20", ">20"),
                     colorpal = pal))



(plt_cereals <- 
    rasterize(country_data %>% filter(basketItem2 == "cereals") %>% 
                group_by(FAO_ID, country) %>% 
                summarise(foodloss_mean = mean(loss_percentage1)) ,
              template_rast_5arcmin, "foodloss_mean")  %>% 
    create_index_map(., tocrs = "ESRI:54030",
                     index_main_title = "(d) Other cereals & pulses",
                     index_label = "Average food loss (%)",
                     breakvals = c(0,2,4,6,8,10,15,20,Inf),
                     breaknames = c("<2", "2-4", "4-6","6-8", "8-10", "10-15", "15-20", ">20"),
                     colorpal = pal))



(plt_fruits <- 
    rasterize(country_data %>% filter(basketItem2 == "fruits") %>% 
                group_by(FAO_ID, country) %>% 
                summarise(foodloss_mean = mean(loss_percentage1)) ,
              template_rast_5arcmin, "foodloss_mean")  %>% 
    create_index_map(., tocrs = "ESRI:54030",
                     index_main_title = "(e) Fruits & vegetables",
                     index_label = "Average food loss (%)",
                     breakvals = c(0,2,4,6,8,10,15,20,Inf),
                     breaknames = c("<2", "2-4", "4-6","6-8", "8-10", "10-15", "15-20", ">20"),
                     colorpal = pal))




(plt_roots <- 
    rasterize(country_data %>% filter(basketItem2 == "roots") %>% 
                group_by(FAO_ID, country) %>% 
                summarise(foodloss_mean = mean(loss_percentage1)) ,
              template_rast_5arcmin, "foodloss_mean")  %>% 
    create_index_map(., tocrs = "ESRI:54030",
                     index_main_title = "(f) Roots, tubers, & oil bearing crops",
                     index_label = "Average food loss (%)",
                     breakvals = c(0,2,4,6,8,10,15,20,Inf),
                     breaknames = c("<2", "2-4", "4-6","6-8", "8-10", "10-15", "15-20", ">20"),
                     colorpal = pal))


(plts_food_losses_commodities <- tmap_arrange(
  plt_wheat, plt_maize, 
  plt_rice, plt_cereals,
  plt_fruits, plt_roots,
  ncol = 2))


# uncomment these!
# tmap_save(plts_food_losses_commodities,
#           "average_food_losses_commodities_rev.png",
#           dpi = 300)
# 
# 
# tmap_save(plts_food_losses_commodities,
#           "average_food_losses_commodities_rev.pdf",
#           dpi = 300)

```



# Figure 2

Average food loss data distribution based on different food supply stages across commodities: farm (a), harvest (b), storage (c), and transport (d). 


```{r}
(plt_farm <- 
    rasterize(country_data %>% filter(food_supply_stage == "Farm") %>% 
    group_by(FAO_ID, country) %>% 
    summarise(foodloss_mean = mean(loss_percentage1)) ,
              template_rast_5arcmin, "foodloss_mean")  %>% 
    create_index_map(., tocrs = "ESRI:54030",
                     index_main_title = "(a) Farm",
                     index_label = "Average food loss (%)",
                     breakvals = c(0,2,4,6,8,10,15,20,Inf),
                     breaknames = c("<2", "2-4", "4-6","6-8", "8-10", "10-15", "15-20", ">20"),
                     colorpal = pal))



(plt_harvest <- 
    rasterize(country_data %>% filter(food_supply_stage == "Harvest") %>% 
    group_by(FAO_ID, country) %>% 
    summarise(foodloss_mean = mean(loss_percentage1)) ,
              template_rast_5arcmin, "foodloss_mean")  %>% 
    create_index_map(., tocrs = "ESRI:54030",
                     index_main_title = "(b) Harvest",
                     index_label = "Average food loss (%)",
                     breakvals = c(0,2,4,6,8,10,15,20,Inf),
                     breaknames = c("<2", "2-4", "4-6","6-8", "8-10", "10-15", "15-20", ">20"),
                     colorpal = pal))



(plt_storage <- 
    rasterize(country_data %>% filter(food_supply_stage == "Storage") %>% 
    group_by(FAO_ID, country) %>% 
    summarise(foodloss_mean = mean(loss_percentage1)) ,
              template_rast_5arcmin, "foodloss_mean")  %>% 
    create_index_map(., tocrs = "ESRI:54030",
                     index_main_title = "(c) Storage",
                     index_label = "Average food loss (%)",
                     breakvals = c(0,2,4,6,8,10,15,20,Inf),
                     breaknames = c("<2", "2-4", "4-6","6-8", "8-10", "10-15", "15-20", ">20"),
                     colorpal = pal))



(plt_transport <- 
    rasterize(country_data %>% filter(food_supply_stage == "Transport") %>% 
    group_by(FAO_ID, country) %>% 
    summarise(foodloss_mean = mean(loss_percentage1)) ,
              template_rast_5arcmin, "foodloss_mean")  %>% 
    create_index_map(., tocrs = "ESRI:54030",
                     index_main_title = "(d) Transport",
                     index_label = "Average food loss (%)",
                     breakvals = c(0,2,4,6,8,10,15,20,Inf),
                     breaknames = c("<2", "2-4", "4-6","6-8", "8-10", "10-15", "15-20", ">20"),
                     colorpal = pal))



(plts_food_losses_supply_stage <- tmap_arrange(
  plt_farm,
  plt_harvest,
  plt_storage, 
  plt_transport, 
  ncol = 2))


# uncomment these!
# tmap_save(plts_food_losses_supply_stage,
#           "average_food_losses_supply_stages_rev.png",
#           dpi = 300)
# 
# 
# tmap_save(plts_food_losses_supply_stage,
#           "average_food_losses_supply_stages_rev.pdf",
#           width = 180, height = 150, units = "mm")

```



# Figure Distributions of different socioeconomic variables

```{r}
# Extract data for explanatory variables, average values for 9 independent variables
explanatory_variables_averages <-
  country_data %>%
  group_by(FAO_ID) %>%
  summarise(gni_mean = mean(gni),
            agriShr_mean = mean(agriShr),
            agriEmp_mean = mean(agriEmp),
            years_of_schooling_mean = mean(msch),
            electricityRural_mean = mean(electricityRural),
            phoneSub_mean = mean(phoneSub),
            polStab_mean = mean(polStab),
            export_mean = mean(export),
            median_incLevel_num = median(as.numeric(factor(incLevel, levels = c("L", "M", "H")))))






# plots

(plt_gni_mean <- rasterize(explanatory_variables_averages, template_rast_5arcmin, "gni_mean") %>% 
  create_index_map(., tocrs = "ESRI:54030",
                   index_main_title = "(a) GNI per capita",
                   index_label = "International $",
                   breakvals = c(0,1000,2000,4000,6000,10000,Inf),
                   breaknames = c("<1000", "1000-2000", "2000-4000","4000-6000", "6000-10000", ">10000"),
                   colorpal = pal))




(plt_agriShr_mean <- rasterize(explanatory_variables_averages, template_rast_5arcmin, "agriShr_mean") %>% 
  create_index_map(., tocrs = "ESRI:54030",
                   index_main_title = "(b) Agriculture shares of GDP",
                   index_label = "% of GDP",
                   breakvals = seq(0, 100, 10),
                   breaknames = c("<10", "10-20", "20-30", "30-40", "40-50",
                                  "50-60", "60-70","70-80", "80-90", "90-100"), 
                   colorpal = pal))




(plt_agriEmp_mean <- rasterize(explanatory_variables_averages, template_rast_5arcmin, "agriEmp_mean") %>% 
  create_index_map(., tocrs = "ESRI:54030",
                   index_main_title = "(c) Employment in agriculture",
                   index_label = "% of total employment",
                   breakvals = seq(0, 100, 10),
                   breaknames = c("<10", "10-20", "20-30", "30-40", "40-50",
                                  "50-60", "60-70","70-80", "80-90", "90-100"), 
                   colorpal = pal))



(plt_years_of_schooling_mean <- rasterize(explanatory_variables_averages, template_rast_5arcmin, "years_of_schooling_mean") %>% 
  create_index_map(., tocrs = "ESRI:54030",
                   index_main_title = "(d) Education",
                   index_label = "Years",
                   breakvals = c(0,2,4,6,8,10,Inf),
                   breaknames = c("<2", "2-4", "4-6", "6-8", "8-10",">10"), 
                   colorpal = pal))



(plt_electricityRural_mean <- rasterize(explanatory_variables_averages, template_rast_5arcmin, "electricityRural_mean") %>% 
  create_index_map(., tocrs = "ESRI:54030",
                   index_main_title = "(e) Access to electricity in rural area",
                   index_label = "% of rural population",
                   breakvals = seq(0, 100, 10),
                   breaknames = c("<10", "10-20", "20-30", "30-40", "40-50",
                                  "50-60", "60-70","70-80", "80-90", "90-100"), 
                   colorpal = pal))


(plt_phoneSub_mean <- rasterize(explanatory_variables_averages, template_rast_5arcmin, "phoneSub_mean") %>% 
  create_index_map(., tocrs = "ESRI:54030",
                   index_main_title = "(f) Mobile cellular subscriptions",
                   index_label = "% of population",
                   breakvals = seq(0, 100, 10),
                   breaknames = c("<10", "10-20", "20-30", "30-40", "40-50",
                                  "50-60", "60-70","70-80", "80-90", "90-100"), 
                   colorpal = pal))



(plt_polStab_mean <- rasterize(explanatory_variables_averages, template_rast_5arcmin, "polStab_mean") %>% 
  "*" (100) %>% 
  create_index_map(., tocrs = "ESRI:54030",
                   index_main_title = "(g) Political stability and absence of violence",
                   index_label = "Index (low value=less stable; high value=more stable)",
                   breakvals = seq(0, 100, 10),
                   breaknames = c("<10", "10-20", "20-30", "30-40", "40-50",
                                  "50-60", "60-70","70-80", "80-90", "90-100"), 
                   colorpal = pal))




(plt_export_mean <- rasterize(explanatory_variables_averages, template_rast_5arcmin, "export_mean") %>% 
  create_index_map(., tocrs = "ESRI:54030",
                   index_main_title = "(h) Export volume index",
                   index_label = "Index (low value=less export; high value=more export)",
                   breakvals = c(0, 50 , 100, 150, 200, 250, 500, 1000, Inf),
                   breaknames = c("<50", "50-100", "100-150", "150-200", "200-250",
                                 "250-500", "500-1000", ">1000"),
                   colorpal = pal))


# Income level
# plt_incLevel <- rasterize(explanatory_variables_averages, template_rast_5arcmin, "median_incLevel_num") %>% 
#   create_index_map(., tocrs = "ESRI:54030",
#                    index_main_title = "(i) Income level",
#                    index_label = "Income level",
#                    breakvals = c(0.5, 1.5, 2.5, 3.5),
#                    breaknames = c("Low", "Middle", "High"),
#                    colorpal = pal)
# 
# plt_incLevel




plts_figureAx <- tmap_arrange(plt_gni_mean, plt_agriShr_mean,
                              plt_agriEmp_mean, plt_years_of_schooling_mean,
                              plt_electricityRural_mean, plt_phoneSub_mean, 
                              plt_polStab_mean, plt_export_mean,
                              ncol = 2) #  plt_incLevel removed
plts_figureAx


# # uncomment these!
# tmap_save(plts_figureAx,
#           filename = "average_explanatory_variables_rev.png",
#           dpi = 300)
# 
# 
# tmap_save(plts_figureAx,
#           filename = "average_explanatory_variables_rev.pdf",
#           width = 180, height = 150, units = "mm")

```


# Table experiment

```{r}
# Extract data for explanatory variables, average values for 9 independent variables
library(sf)
library(plotrix)

df_descriptive_using_countrydata <-
  country_data %>%
  st_drop_geometry() %>% 
  group_by(incLevel) %>%
  summarise(foodloss_mean = mean(loss_percentage1),
    gni_mean = mean(gni),
    gni_se = std.error(gni),
    agriShr_mean = mean(agriShr),
    agriShr_se = std.error(agriShr),
    agriEmp_mean = mean(agriEmp),
    agriEmp_sd =  std.error(agriEmp),
    years_of_schooling_mean = mean(msch),
    years_of_schooling_se = std.error(msch),
    electricityRural_mean = mean(electricityRural),
    electricityRural_se =  std.error(electricityRural),
    phoneSub_mean = mean(phoneSub),
    phoneSub_se = std.error(phoneSub),
    polStab_mean = mean(polStab),
    polStab_se =  std.error(polStab),
    export_mean = mean(export),
    export_se =  std.error(export))


```


#
